Speed and strength of an epidemic intervention

### 2020-10-15 response.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}

\usepackage{color}

\usepackage{xspace}

\usepackage{lmodern}
\usepackage{amssymb,amsmath}

\usepackage[pdfencoding=auto, psdextra]{hyperref}

\usepackage{natbib}
\bibliographystyle{chicago}

\newcommand{\eref}[1]{Eq.~(\ref{eq:#1})}
\newcommand{\fref}[1]{Fig.~\ref{fig:#1}}

\newcommand{\rr}{\ensuremath{{r}}}
\newcommand{\rR}{\mbox{$r$--$\cal R$}}
\newcommand{\RR}{\ensuremath{{\cal R}}}
\newcommand{\RRhat}{\ensuremath{{\hat \cal R}}}
\newcommand{\Rx}[1]{\ensuremath{{\cal R}_{#1}}} 
\newcommand{\Ro}{\ensuremath{{\mathcal R}_{0}}\xspace}
\newcommand{\Rs}{\Rx{\mathrm{s}}}
\newcommand{\Rpool}{\ensuremath{{\mathcal R}_{\textrm{\tiny{pool}}}}\xspace}
\newcommand{\Reff}{\Rx{\mathit{eff}}}
\newcommand{\Tc}{\ensuremath{C}}

\newcommand{\dd}[1]{\ensuremath{\, \mathrm{d}#1}}
\newcommand{\dtau}{\dd{\tau}}
\newcommand{\dx}{\dd{x}}
\newcommand{\dsigma}{\dd{\sigma}}

\newcommand{\rev}{\subsection*}
\newcommand{\revtext}{\textsf}
\setlength{\parskip}{\baselineskip}
\setlength{\parindent}{0em}

\newcommand{\comment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
\newcommand{\jd}[1]{\comment{cyan}{JD}{#1}}
\newcommand{\swp}[1]{\comment{magenta}{SWP}{#1}}
\newcommand{\dc}[1]{\comment{blue}{DC}{#1}}
\newcommand{\jsw}[1]{\comment{green}{JSW}{#1}}
\newcommand{\hotcomment}[1]{\comment{red}{HOT}{#1}}

\newcommand{\psymp}{\ensuremath{p}} %% primary symptom time
\newcommand{\ssymp}{\ensuremath{s}} %% secondary symptom time
\newcommand{\pinf}{\ensuremath{\alpha_1}} %% primary infection time
\newcommand{\sinf}{\ensuremath{\alpha_2}} %% secondary infection time

\newcommand{\psize}{{\mathcal P}} %% primary cohort size
\newcommand{\ssize}{{\mathcal S}} %% secondary cohort size

\newcommand{\gtime}{\tau_{\rm g}} %% generation interval
\newcommand{\gdist}{g} %% generation-interval distribution
\newcommand{\idist}{\ell} %% incubation period distribution

\newcommand{\total}{{\mathcal T}} %% total number of serial intervals

\begin{document}

\noindent Dear Editor:

Thank you for the chance to revise and resubmit our manuscript. 
We have made major revisions to our manuscript to address the reviewers' comments.
We have also added a new section and discuss how presymptomatic transmission can affect the efficacy of COVID-19 interventions.
Below please find our detailed responses to reviewers.

\rev{Reviewer \#1}

\revtext{One additional reference that might be worth mentioning is O. Diekmann,
H. Heesterbeek, and H. Metz,The legacy of Kermack and McKendrick, in
Epidemic Models, Their Structure and Relation to Data (D. Mollinson,
ed.) (1995), pp.95-115.}

Thank you. We now cite this article as an early mentioner of the
little-r threshold.

\rev{Reviewer \#2}

\revtext{In this work the authors propose that ``strength'' and ``speed'' are
complementary frameworks for understanding interventions in infectious
disease models. This is an interesting and appealing idea based on
simple re-arrangements of well-known equations for exponentially growing
epidemics. As the authors themselves note, comparing growth rates r
rather than reproductive numbers R is already well known in ecology.}

\revtext{In my view this paper could bring these ideas to light in infectious
disease more thoroughly and compellingly with an expanded application
section, a section on the relationship to data and/or estimation from
data, improved and expanded section on interventions and the distinct
views provided by (and usefulness of) the two distinct frameworks,
and/or with some extension of the results to population dynamics that
differ from constant rate exponential growth or decay. As it stands it
feels quite minimal, though I appreciate the appeal of the conceptual
point.}

Thank you for this feedback. We agree that these ideas are similar to
ideas that have been discussed in the ecological context, but also feel
that this is a perspective sorely lacking in the disease-modeling world.

We have expanded the
application section with examples from COVID-19, providing a qualitative
understanding on how uncertainties in the amount of presymptomatic
transmission can affect conclusions about the efficacy of intervention.
In the beginning of the current pandemic, there was much discussion on
how presymptomatic transmission could make intervention more
difficult---however, some of these conclusions were based on the
assumption that \RR\ is constant, rather than \rr\ (e.g., Hellewell et
al.~2020. ``Feasibility of controlling COVID-19 outbreaks by isolation
of cases and contacts.'' \emph{ Lancet Global Health}). In reality, given
observed \rr, more presymptomatic transmission reduces \RR, meaning that
strength-based interventions will be easier than previously thought. To
our mind, these examples further underline the absence of \rr-based
thinking in the disease-modeling world. While we appreciate the
reviewer's concern that a more data-driven approach could provide
further insight, we are worried that details may take away general ideas
behind speed- and strength-based paradigms. We feel that it is important
to provide a conceptual framework at this stage to a broader audience.
We also address more explicitly the link between the parameters of the
initial phase and the prospects for long-term control. We feel that
issues dealing directly with non-exponential disease dynamics are
complicated enough for another paper.

\revtext{Near equation (9) the authors need to make a more explicit link with
survival analysis so that we are clear on why we multiply K by survival
functions. Survival functions multiply in the sense that if there are
two hazards at rates lambda1 and lambda2 then the survival function is
S(t) = S1(t) S2(t) . But K(tau) is proportional to the density g(tau),
and is not a survival function. What is the interpretation of the
density times the survival function? Why would the interventions not
change the density directly, so that g becomes ghat (and then this would
impact the hazard rate and survival function via the usual relationship
$S(t) = \int_t ^\infty ( \hat{g} (x) dx )  )$ ?}

We have expanded (9) into two steps and attempted to explain the logic clearly:

``Now imagine an idealized intervention that reduces transmission at a constant hazard rate $\phi$ across the disease generation (Fig. 1B), for example, by identifying and isolating infectious individuals.
We then have:
\begin{equation}
	\hat K(\tau) = K(\tau)\exp(-\phi\tau)
\end{equation}
The interpretation is that average infectiousness for under this control regime is the product of the original infectiousness $K(\tau)$ (at age of infection $\tau$) and the probability $\exp(-\phi\tau)$ of escaping the hazard of control up to that time.

Substituting (8):
\begin{equation}
	\hat K(\tau) = K(\tau)\exp(-\phi\tau) = b(\tau)\exp((r-\phi)\tau)
\end{equation}
Since $b$ is a distribution (which integrates to 1), the reduction needed to prevent invasion (or to eliminate disease) is exactly $\phi=r$. 
We call $\phi$ the ``speed'' of the intervention; transmission is interrupted when the speed of the intervention is faster than the speed of spread.''

\revtext{
In practice what is the need for equations (6) and (12)? Both
essentially re-write the other equations; are these averages that can be
linked to estimates from data, for example?
}

We think that the value of (6) and (12 -- now 13) is conceptual. We have
added explanations following these derivations. In retrospect, we
believe that S2.2 and S2.3 were written in too ``mathematical'' a
style: we got a bit carried away with the formal parallelism. The new
version is still short, but engages more with the biological concepts.

We note that in practice 6 and 12 are often implicit in mechanistic models. We now make this point in the MS near the end of both S2.2 and S2.3:

``
We note that intervention function $L$ and the strength of intervention $\theta$ need not be calculated explicitly in many contexts: they can usefully be thought of as abstractions of existing modeling practices.
Modelers typically rely on mechanistic models (often based on ordinary differential equations) to model disease spread and evaluate intervention effects.
By doing so, they make implicit assumptions about the shape of $L$ and therefore $\theta$.
''

``Like intervention strength $\theta$, intervention speed $\phi$ is also an abstraction --- that is, the mechanistic models of interventions make implicit assumptions about the shape of the hazard rate $h$ and therefore $\phi$.''

\revtext{
Presumably both frameworks give a decline when the growth rate is
negative and growth when it is positive. Therefore they would never
contradict each other as to whether an intervention was sufficient or
not - is this correct? This should be made explicit.
}

This is correct, and we have worked to be more explicit. 
\swp{Where? Can we get a quote?}

\revtext{
It would also be good to more explicitly state and discuss what the
differences are. I realize that this is in the discussion and in some
sense in Figure 4, but I found the discussion section vague,
particularly since the analysis is only relevant for the case of
exponential growth. In particular, where would the two frameworks give
different estimates of the uncertainty surrounding the benefits of an
intervention, or the comparison between two interventions?
}

We never expect to get a different final answer from the two formalisms if they are both correctly applied;
they are two ways of doing essentially the same calculation. 
We have tried to be more clear about this as mentioned above. 
We do expect in some cases to be able to use one formalism or another to more clearly illuminate the key issues and quantities determining the detailed answer, as we try to explain in our HIV example and in our new COVID example.

Keeping the two paradigms in mind, however, can help to do the analysis correctly.
In particular, in emerging epidemics, researchers often vary assumptions about parameters while holding \RR\ constant, even in cases where it would be more consistent with available information to hold \rr\ constant.
We now discuss this in the paper:

``Thinking explicitly about the two perspectives can also reduce confusion. Because of the dominance of the strength paradigm, researchers often explore different scenarios while holding \RR\ fixed. Fixed \RR\ is in fact a good default assumption for many endemic diseases. For invading diseases, however, \rr\ is likely to be better constrained by data than \RR. In this case, comparing scenarios while holding \RR\ fixed creates a bias that makes scenarios with faster transmission at the individual level \swp{(i.e., higher proportion of early transmission)} look relatively more dangerous (since these scenarios will have \rr\ faster than the observed value). \swp{I think we can remove the parenthesis in the last sentence}''

\revtext{
On a minor note, since condoms are ultimately an individual choice
applied to individual (potential) transmission events, why are they a
population-level intervention whereas test and treat is deemed
individual? More details on why the two frameworks give different
intuition or analytical capacity would be good.
}

Thank you for this. We were using individual to mean targeted at
infectious individuals, and population to mean targeted at everyone. We
have now dropped these shortcuts, and we simply explain this directly:

``We expect the strength-based framework to be clearer for intervention strategies that target the general population, like condom use, or susceptible people, like prophylaxis.''

We have also added more discussion of how frameworks differ, both in examples and in Discussion.

\revtext{In Figures 3 and 4, the black lines are not clear to me -- the strength
is defined as R/Rhat and the speed as rhat -r . So what is the strength
(or speed) of ``the epidemic'' (black lines) - what's Rhat and what's R?
It might be good to have some prevalence curves}

We define strength and speed of the epidemic in the introduction, and
compare these to strength and speed of intervention efforts. We now
reiterate these points (including in the figure captions).

\revtext{
Minor quibble: equation (4) has L described in words as the ``average
proportional reduction'' but where the reduction is 75\%, in fact the
denominator is 4: $(1 / (1-0.75) )$ -- rephrase.
}

Done, thank you.

\end{document}

### 2020-10-15 speedstrength.tex
some more edits

Here, we argue that the dominance of \RR\ over $r$ in the disease-dynamics literature is excessive: $r$ has been under-rated as a metric. Like \RR,  $r$ can serve as a threshold, and we show that it can also provide a useful metric for difficulty of elimination (cf.~\cite{ferretti2020quantifying}). 
We first generalize the idea that \RR\ measures the difficulty of elimination by showing we can measure an intervention's ``strength'' on the same scale as the reproductive number. 
We then show that we can likewise measure an intervention's ``speed'' on the same scale as the growth rate.
Thus, there is a dual sense in which \rr\ also measures difficulty of elimination. 
We argue that the primacy of $\RR$ over $r$ is partly due to history, and discuss cases where strength provides the better framing for practical disease questions and cases where speed does.

@@

Since we are interested in invasion and control, we will generally neglect changes in $K(\tau, t)$ through time.
Thus, we will assume $K(\tau, t) \equiv K(\tau)$.
Importantly, this means we are neglecting changes in susceptible proportion through time: that is, assuming that $S(t) \approx S(0)$.
Under this assumption, the renewal equation is equivalent to the Von Foerster equations (see e.g. \cite{fraser2004factors}).
Note that this focus does not mean we are only interested in the initial period of exponential growth: the ability of a disease to increase under conditions characteristic of a naive population is commonly used as a criterion for whether it can be eliminated under general conditions \citep{anderson1985vaccination,fraser2004factors}.

@@

We define the strength of the intervention $L$ to be $\theta = \RR/\Rhat$. 
It is then straightforward to show that $\theta$ is the harmonic mean of $L(\tau)$ weighted by the intrinsic generation-interval distribution:
\begin{equation}
	\theta = 1/\langle 1/L(\tau) \rangle_{g(\tau)}.
	\eqlab{strengthMean}
\end{equation}
Again, we have that the disease cannot spread when $\theta \geq \RR$. 
In other words, $\theta$ generalizes the well-known idea that \RR\ is a metric for how hard a disease is to eliminate: we can measure the ``strength'' of a control measure, this must exceed the strength of the disease (\RR) to achieve elimination.

@@

The above decomposition generalizes the argument that \RR\ is the key parameter in evaluating whether a disease can be controlled -- one of the main foundations of historical primacy of \RR. But we can in fact do an analogous decomposition based on speed, and place \rr\ in a similar role.

@@

We generalize this idea by allowing the hazard rate $h(\tau)$ at which $K$ is reduced to vary through time, thus:
\begin{equation}
	\hat K(\tau) = K(\tau) \exp\left(-\int_0^\tau h(\sigma) d\sigma\right)
\end{equation}
The associated post-intervention epidemic speed $\hat r$ is given by:
\begin{equation}
	1 = \int \hat K(\tau) \exp(-\hat r\tau) d\tau.	
\end{equation}
We define the speed of a general intervention to be $\phi = r - \hat r$
We can then show that $\phi$ is a (sort of) mean satisfying the implicit equation:
\begin{equation}
	1 = \left\langle \frac{\exp(\phi \tau) }{\exp\left(\int_0^\tau h(\sigma) d\sigma\right)} \right\rangle_{b(\tau)}
	\eqlab{speedMean}
\end{equation}
Specifically, the speed $\phi$ is a mean of the hazard $h$ in the sense that an increase (or decrease) in $h$ produces the same sign of change in $\phi$, and if $h$ is constant across the generation then $\phi=h$.
Like intervention strength $\theta$, intervention speed $\phi$ is also an abstraction -- that is, the mechanistic models of interventions make implicit assumptions about the shape of the hazard rate $h$ and therefore $\phi$.

@@

Though there is a clear intuition for why both strengths increase as early transmission goes down, the speed paradigm provides insight into why these two increases are so close to parallel.
The estimated epidemic speed depends only on the observed growth rate -- it does not change if we change our assumption about the proportion of early transmission.
For the test-and-treat intervention, the effective intervention speed also stays relatively constant (\figref{test}D), in part because we have (plausibly) assumed that hazard stays relatively constant for a few key months, and in part because the backward generation-interval distributions for different scenarios are relatively similar (Supplemental figure).
The effective intervention speed increases slightly as proportion of early transmission increases because the subpopulation that the intervention fails to reach become relatively more important if late transmission is more important.
Thus, the speed paradigm provides an intuitive underpinning for the originally surprising result of \cite{eaton2014proportion}: the effectiveness of test-and-treat interventions should not depend much on the proportion of early transmission.

@@

There has been a great deal of discussion of the importance of pre-symptomatic transmission of COVID-19. 
Pre-symptomatic transmission is likely to be hard to detect, and therefore hard to prevent. 
Thus, it might be supposed that an increase in the estimated importance of pre-symptomatic transmission would lead to an increase in the estimated difficulty of control.

A generation-interval perspective \cite{park2019practical} can be used to challenge this view. 
More early transmission, for a given value of \rr, means {shorter generation intervals and therefore} less transmission per individual -- that is, a lower value of \RR. Thus, although earlier transmission could make intervention less effective, it also means that less intervention may be necessary. For interventions like lockdowns, which are generally assumed affect everyone roughly equally, the strength-based perspective thus gives a clear (and somewhat surprising) answer: more early transmission means we will conclude that non-targeted interventions are \emph{more} effective.

@@

We also discussed the question of uncertainties introduced by the unknown magnitude of pre-symptomatic transmission of COVID-19. We showed that the strength-based paradigm provides clear insight into how this uncertainty affects interventions targeted to the general population, while the speed-based paradigm is a clearer way to think about interventions targeted to infected people. We concluded that increasing pre-symptomatic transmission increases estimated effectiveness of contact-based interventions and decreases estimated effectiveness of symptom-based interventions. In hindsight, these conclusions are consistent with common sense, but in practice the speed-based framework provides a clear way to think about these questions.

@@

Thinking explicitly about the two perspectives can also reduce confusion. Because of the dominance of the strength paradigm, researchers often explore different scenarios while holding \RR\ fixed. Fixed \RR\ is in fact a good default assumption for many endemic diseases. For invading diseases, however, \rr\ is likely to be better constrained by data than \RR. In this case, comparing scenarios while holding \RR\ fixed creates a bias that makes scenarios with faster transmission at the individual level look relatively more dangerous (since these scenarios will have \rr\ faster than the observed value) \citep{eaton2014proportion, powers2014impact, park2019practical}.

For interventions, we expect the speed-based framework to be clearer for evaluating intervention strategies that target infected people, like test-and-treat for HIV \citep{granich2009universal}, or contact-tracing and quarantine for COVID-19 \citep{hellewell2020feasibility}; we expect the strength-based framework to be clearer for intervention strategies that target the general population, like condom use, or susceptible people, like prophylaxis.
In other cases, such as real-time rollout of vaccines during an outbreak, both strength and speed approaches might be similarly uncertain because the result depends both on the speed of the rollout and the (strength-like) final coverage \citep{shah2018mumps}.

@@

The importance of speed-based perspectives are still rarely recognized in the case of infectious disease, however. 
Responses to the 2014 Ebola Outbreak in West Africa and the recent COVID-19 outbreak show an over-emphasis on strength at the expense of speed:   
during the early phases of both outbreaks, many disease modelers tried to estimate $\mathcal R_0$ but overlooked $r$.
For example, only 1 out of 7 preliminary analyses of the COVID-19 outbreak that were published as preprints between January 23--26, 2020 reported the doubling time of an epidemic \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}. \swp{does this feel weak now? Maybe need something like ``Many subsequent studies then relied on strength-based frameworks to evaluate the efficacy of speed-like interventions, such as contact tracing, with a few exceptions'' with citations?}
\jd{It would be great if you could do this.}
We suggest that infectious disease modelers should be aware of the complementarity of these two frameworks when analyzing disease outbreaks.

>>>>

Here, we argue that the dominance of \RR\ over $r$ in the disease-dynamics literature is excessive: $r$ has been under-rated as a metric. Like \RR,  $r$ can serve as a threshold, and we show that it can also provide a useful metric for difficulty of elimination (cf.~\cite{ferretti2020quantifying}). 
We first generalize the idea that \RR\ measures the difficulty of elimination by showing we can measure an intervention's ``strength'' on the same scale as the reproductive number. 
We then show that we can likewise measure an intervention's ``speed'' on the same scale as the growth rate.
Thus, there is a dual sense in which \rr\ also measures difficulty of elimination. 
We argue that the primacy of $\RR$ over $r$ is partly due to history, and discuss cases where strength provides the better framing for practical disease questions and cases where speed does.
\swp{maybe this is where it might be better to say "speed provides the better framing ..."?}

@@

Since we are interested in invasion and control, we will generally neglect changes in $K(\tau, t)$ through time.
Thus, we will assume $K(\tau, t) \equiv K(\tau)$.
Importantly, this means we are neglecting changes in susceptible proportion through time: that is, assuming that $S(t) \approx S(0)$.
Under this assumption, the renewal equation is equivalent to the Von Foerster equations (see e.g. \cite{fraser2004factors}).
Note that this focus does not mean we are only interested in the initial period of exponential growth: the ability of a disease to increase under conditions characteristic of a naive population is commonly used as a criterion for whether it can be eliminated under general conditions \citep{anderson1985vaccination,fraser2004factors}.
\swp{I feel like there's too much ``something, ...'' phrasing. How about the following?} 
\swp{
Since we are interested in invasion and control, we will generally neglect changes in $K(\tau, t)$ through time and assume $K(\tau, t) \equiv K(\tau)$.
Importantly, this means we are neglecting changes in susceptible proportion through time: $S(t) \approx S(0)$.
Under this assumption, the renewal equation is equivalent to the Von Foerster equations (see e.g. \cite{fraser2004factors}).
However, this focus does not mean we are only interested in the initial period of exponential growth: the ability of a disease to spread under conditions characteristic of a naive population is commonly used as a criterion for whether it can be eliminated under general conditions \citep{anderson1985vaccination,fraser2004factors}.
}
\swp{``conditions characteristic of a naive population'' doesn't really make sense...}

@@

We define the strength of the intervention $L$ to be $\theta = \RR/\Rhat$. 
It is then straightforward to show that $\theta$ is the harmonic mean of $L(\tau)$ weighted by the intrinsic generation-interval distribution:
\begin{equation}
	\theta = 1/\langle 1/L(\tau) \rangle_{g(\tau)}.
	\eqlab{strengthMean}
\end{equation}
Again, we have that the disease cannot spread when $\theta \geq \RR$. 
In other words, $\theta$ generalizes the well-known idea that \RR\ is a metric for how hard a disease is to eliminate: we can measure the ``strength'' of a control measure, this must exceed the strength of the disease (\RR) to achieve elimination.
\swp{The last two/three sentences feel repetitive.}

@@

The above decomposition generalizes the argument that \RR\ is the key parameter in evaluating whether a disease can be controlled --- one of the main foundations of historical primacy of \RR. But we can in fact do an analogous decomposition based on speed, and place \rr\ in a similar role.

@@

We generalize this idea by allowing the hazard rate $h(\tau)$ at which $K$ is reduced to vary through time, thus:
\begin{equation}
	\hat K(\tau) = K(\tau) \exp\left(-\int_0^\tau h(\sigma) d\sigma\right)
\end{equation}
The associated post-intervention epidemic speed $\hat r$ is given by:
\begin{equation}
	1 = \int \hat K(\tau) \exp(-\hat r\tau) d\tau.	
\end{equation}
We define the speed of a general intervention to be $\phi = r - \hat r$
We can then show that $\phi$ is a (sort of) mean satisfying the implicit equation:
\begin{equation}
	1 = \left\langle \frac{\exp(\phi \tau) }{\exp\left(\int_0^\tau h(\sigma) d\sigma\right)} \right\rangle_{b(\tau)}
	\eqlab{speedMean}
\end{equation}
Specifically, the speed $\phi$ is a mean of the hazard $h$ in the sense that an increase (or decrease) in $h$ produces the same sign of change in $\phi$, and if $h$ is constant across the generation then $\phi=h$.
Like intervention strength $\theta$, intervention speed $\phi$ is also an abstraction --- that is, the mechanistic models of interventions make implicit assumptions about the shape of the hazard rate $h$ and therefore $\phi$.

@@

Though there is a clear intuition for why both strengths increase as early transmission goes down, the speed paradigm provides insight into why these two increases are so close to parallel.
The estimated epidemic speed depends only on the observed growth rate --- it does not change if we change our assumption about the proportion of early transmission.
For the test-and-treat intervention, the effective intervention speed also stays relatively constant (\figref{test}D), in part because we have (plausibly) assumed that hazard stays relatively constant for a few key months, and in part because the backward generation-interval distributions for different scenarios are relatively similar (Supplemental figure).
The effective intervention speed increases slightly as proportion of early transmission increases because the subpopulation that the intervention fails to reach become relatively more important if late transmission is more important.
Thus, the speed paradigm provides an intuitive underpinning for the originally surprising result of \cite{eaton2014proportion}: the effectiveness of test-and-treat interventions should not depend much on the proportion of early transmission.

@@

There has been a great deal on discussion of the importance of pre-symptomatic transmission of COVID-19 \cite{ferretti2020quantifying, he2020temporal, hellewell2020feasibility}. 
Pre-symptomatic transmission is likely to be hard to detect, and therefore hard to prevent. 
Thus, it might be supposed that an increase in the estimated importance of pre-symptomatic transmission would lead to an increase in the estimated difficulty of control.

A generation-interval perspective \cite{park2019practical} can be used to challenge this view. 
More early transmission, for a given value of \swp{observed} \rr, means {shorter generation intervals and therefore} less transmission per individual --- that is, a lower value of \RR. Thus, although earlier transmission could make intervention less effective, it also means that less intervention may be necessary. For interventions like lockdowns, which are generally assumed \swp{to} affect everyone roughly equally, the strength-based perspective thus gives a clear (and somewhat surprising) answer: more early transmission means we will conclude that non-targeted interventions are \emph{more} effective.

@@

We also discussed the question of uncertainties introduced by the unknown magnitude of pre-symptomatic transmission of COVID-19. We showed that the strength-based paradigm provides clear insight into how this uncertainty affects interventions targeted to the general population, while the speed-based paradigm is a clearer way to think about interventions targeted to infected people. We concluded that \swp{``a higher estimate of'' instead of ``increasing''?}increasing pre-symptomatic transmission increases estimated effectiveness of contact-based interventions and decreases estimated effectiveness of symptom-based interventions. In hindsight, these conclusions are consistent with common sense, but in practice the speed-based framework provides a clear way to think about these questions.

@@

Thinking explicitly about the two perspectives can also reduce confusion. Because of the dominance of the strength paradigm, researchers often explore different scenarios while holding \RR\ fixed. Fixed \RR\ is in fact a good default assumption for many endemic diseases. For invading diseases, however, \rr\ is likely to be better constrained by data than \RR. In this case, comparing scenarios while holding \RR\ fixed creates a bias that makes scenarios with faster transmission at the individual level \swp{(i.e., higher proportion of early transmission)} look relatively more dangerous (since these scenarios will have \rr\ faster than the observed value) \citep{eaton2014proportion, powers2014impact, park2019practical}. \swp{I think we can remove the parenthesis in the last sentence}

For interventions, we expect the speed-based framework to be clearer for evaluating intervention strategies that target infected people, like test-and-treat for HIV \citep{granich2009universal}, or contact-tracing and quarantine for COVID-19 \citep{hellewell2020feasibility}. 
We expect the strength-based framework to be clearer for intervention strategies that target the general population, like condom use, or susceptible people, like prophylaxis.
In other cases, such as real-time rollout of vaccines during an outbreak, both strength and speed approaches might be similarly uncertain because the result depends both on the speed of the rollout and the (strength-like) final coverage \citep{shah2018mumps}.

@@

The importance of speed-based perspectives are still rarely recognized in the case of infectious disease, however. 
Responses to the 2014 Ebola Outbreak in West Africa and the recent COVID-19 outbreak show an over-emphasis on strength at the expense of speed:   
during the early phases of both outbreaks, many disease modelers tried to estimate $\mathcal R_0$ but overlooked $r$.
For example, only 1 out of 7 preliminary analyses of the COVID-19 outbreak that were published as preprints between January 23--26, 2020 reported the doubling time of an epidemic \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
Subsequent studies then relied on strength-based frameworks to evaluate the efficacy of speed-like interventions, such as contact tracing \cite{hellewell2020feasibility,kretzschmar2020impact,kucharski2020contact}, with a few exceptions \cite{ferretti2020quantifying}.
\swp{How is this? Don't want to overwhelm with citations. Seems good enough for me.}
We suggest that infectious disease modelers should be aware of the complementarity of these two frameworks when analyzing disease outbreaks.
